---
title: "Chess Tournament ETL"
author: "Angel Claudio"
output:
  html_document:
    df_print: paged
  pdf_document: default
editor_options:
  chunk_output_type: inline
---

```{r setup, message=F}
library(tidyverse)
library(DT)
```

**First step is to extract the data.**

1. Skip first 4 rows (not using them)
2. Give a column name to call on and reject headers
3. Remove all rows using "----.." by leveraging modulus

```{r extract, message=F}
chess_tournament <- read_csv(
    "https://raw.githubusercontent.com/AngelClaudio/DataSources/master/tournamentinfo.txt",
        col_name = ("raw"), skip = 4) %>% 
         filter(row_number() %% 3 != 0
    )
```

**Second step is from the raw data, extract to vectors using regex patterns. There will be a split of two sets of data, the top rows of the players, and the bottom row of the players so they match with the ordinality of the IDs.**

1. Start with vector for Player IDs to use for reference later.
2. Create a vector for Player Names
3. Create a vector for States
4. Create a vector for Points
5. Create a vector for Pre-rating
6. Merge the vectors into a data frame

```{r extract-to-vectors}
player_id = str_extract(chess_tournament$raw,"^\\d+")
player_id = player_id[!is.na(player_id)]

#Update Raw data, removing IDs to make proceeding regex needs simpler
chess_tournament <- data.frame(str_replace(chess_tournament$raw,"^\\d+\\s\\|\\s",""))

player_name_rows <- chess_tournament %>% filter(row_number() %% 2 != 0) 
names(player_name_rows)[1] <- "col1"
player_name = trimws(str_extract(player_name_rows$col1, "^\\D+ [^\\|]"))

state_rows <- chess_tournament %>% filter(row_number() %% 2 == 0) 
names(state_rows)[1] <- "col1"
state_name = str_extract(state_rows$col1, "[[:alpha:]]{2}")

points = str_extract(player_name_rows$col1, "\\d.\\d")
pre_rating = as.integer(str_replace_all(str_extract(state_rows$col1, ": \\s?\\d{3,4}"), ":\\s",""))

chess_tournament <- data.frame(player_id, player_name, state_name, points, pre_rating)

datatable(chess_tournament[2:5], extensions = c('FixedColumns',"FixedHeader"),
          options = list(scrollX = TRUE, 
                         paging=TRUE,
                         fixedHeader=TRUE))
```


**Third step is to get average of all the player's opponents into a variable**

1. Create a vector of all the opponent IDs per player
2. Instantiate a vector that will be used to populate the averages
3. Using a for loop, iterate through all the sets of opponent ids
4. Using the reference IDs of the main set, pull their ratings to use for a mean
4. Take the now populated vector from step 2 and merge with previously created data frame

```{r mean-ratings, warning=F}
players_opponent_ids = str_extract_all(player_name_rows$col1, "[[:upper:]]\\s+\\d+\\|")
players_opponent_ids = str_extract_all(players_opponent_ids, "\\d+")

average_opponent_rating <- vector()

for (ids in players_opponent_ids) {
    opponent_ratings <- filter(chess_tournament, player_id %in% ids)
    average_opponent_rating <- c(average_opponent_rating,  mean(opponent_ratings$pre_rating))
}

chess_tournament$average_opponent_rating <- average_opponent_rating

datatable(chess_tournament[2:6], extensions = c('FixedColumns',"FixedHeader"),
          options = list(scrollX = TRUE, 
                         paging=TRUE,
                         fixedHeader=TRUE))
```

**Fourth Step is to load the new data by writing a CSV (commented out since no visualization)**
```{r}
#Write to CSV file
#write.csv(chess_tournament,"chess_tournament.csv")
```

