---
title: "Week3"
author: "Angel Claudio"
date: "2/10/2020"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
library(stringr)
```


**1. Using the 173 majors listed in fivethirtyeight.com’s College Majors dataset [https://fivethirtyeight.com/features/the-economic-guide-to-picking-a-college-major/], provide code that identifies the majors that contain either "DATA" or "STATISTICS"**

## Answer 1
```{r filter-grepl-demo, message=F}
majors <- read_csv("https://raw.githubusercontent.com/fivethirtyeight/data/master/college-majors/majors-list.csv")

(filter(majors, grepl("DATA|STATISTICS", Major)))
```

**2 Write code that transforms the data below:**
[1] "bell pepper"  "bilberry"     "blackberry"   "blood orange"

[5] "blueberry"    "cantaloupe"   "chili pepper" "cloudberry"  

[9] "elderberry"   "lime"         "lychee"       "mulberry"    

[13] "olive"        "salal berry"

Into a format like this:

c("bell pepper", "bilberry", "blackberry", "blood orange", "blueberry", "cantaloupe", "chili pepper", "cloudberry", "elderberry", "lime", "lychee", "mulberry", "olive", "salal berry")

## Answer 2
```{r collapse-demo}
(c("bell pepper", "bilberry", "blackberry", "blood orange", "blueberry", "cantaloupe", "chili pepper", "cloudberry", "elderberry","lime", "lychee", "mulberry", "olive", "salal berry") %>%
    str_c(collapse=", "))
```

**#3 Describe, in words, what these expressions will match:**

## Answer 3
* (.)\1\1
1. A one character group, repeated, and then repeated again:
```{r}
str_view("aaa", "(.)\\1\\1")
```

* "(.)(.)\\2\\1"
2. Two one character groups, repeated but in reverse order:
```{r}
str_view("goog", "(.)(.)\\2\\1")
```

* (..)\1
3. A two character group repeated:
```{r}
str_view("xyxy", "(..)\\1")
```

* "(.).\\1.\\1"
4. A onc character group with a character repeated, then just group repeated again:
```{r}
str_view("gogog", "(.).\\1.\\1")

```

* "(.)(.)(.).*\\3\\2\\1"
5. Three one character groups repeated but in reverse order:
```{r}
str_view("123321", "(.)(.)(.).*\\3\\2\\1")

```

#4 Construct regular expressions to match words that:

Start and end with the same character.
Contain a repeated pair of letters (e.g. "church" contains "ch" repeated twice.)
Contain one letter repeated in at least three places (e.g. "eleven" contains three "e"s.)
